Type Refinements for the Working Class

### 2015-12-19 paper.tex
comment out stuff pending fix

In general, we must explain typehood for parametric refinements
$\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
transformation
$\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
is a unique $\psi$ such that the following diagram commutes, where
$\lambda(\chi)$ is the exponential transpose of $\chi$:
\[
  \IsValueType{\Upsilon}{\phi}{\tau}
  \triangleq
  \begin{tikzcd}[sep=huge]
    \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
      \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
      \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
    \IMode{\Omega}\\
    \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
      \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
  \end{tikzcd}
\]

Contrary to Martin-L\"of, however, we wish to consider a full range of
\emph{computational types}, including those which are not necessarily defined
by their values; a general apparatus for computation-respecting refinements will
allow us to define, for instance, a type for partial functions.

Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
$\IsType{\Upsilon}{\phi}{\tau}$ just when for any
\begin{tikzcd}[cramped]
  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd}%
, and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
$\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
Stated internally in the topos, the meaning of parametric typehood for
$\phi$ is perhaps a bit more clear:
\[
  \Define{
    \IsType{-}{\phi}{\tau}
  }{
    \prod_{L,M,N\in\Exprs_\tau}
    {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
  }
\]

>>>>

%% In general, we must explain typehood for parametric refinements
%% $\Refines{\Upsilon}{\phi}{\tau}$ as well, however; recall from
%% Remark~\ref{rem:internal} that the definition of $\phi$ consists in a natural
%% transformation
%% $\Of{\phi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Exprs_\tau}}}$.
%% Therefore, we say that $\IsValueType{\Upsilon}{\phi}{\tau}$ obtains when there
%% is a unique $\psi$ such that the following diagram commutes, where
%% $\lambda(\chi)$ is the exponential transpose of $\chi$:
%% \[
%%   \IsValueType{\Upsilon}{\phi}{\tau}
%%   \triangleq
%%   \begin{tikzcd}[sep=huge]
%%     \IMode{\Yoneda{\Upsilon}\times\Values_\tau^2}
%%       \arrow[r, dotted, "\OMode{\lambda(\psi)}"]
%%       \arrow[d, swap, hook, "\Pair{\ArrId{}}{\IMode{\iota}}"] &
%%     \IMode{\Omega}\\
%%     \IMode{\Yoneda{\Upsilon}\times\Exprs_\tau^2}
%%       \arrow[swap, ur, "\IMode{\lambda(\phi)}"]
%%   \end{tikzcd}
%% \]
%%
%% Equivalently, and perhaps more simply, we can express value typehood internally
%% in our topos as a presheaf:
%% \[
%%   \Define{
%%     \IsValueType{-}{\phi}{\tau}
%%   }{
%%     \sum_{\psi\in\Pow{\Values_\tau^2}}
%%     \prod_{M,N\in\Values_\tau}
%%     {\phi(M,N)}^{\psi(M,N)}
%%   }
%% \]
%%
%% Contrary to Martin-L\"of, however, we wish to consider a full range of
%% \emph{computational types}, including those which are not necessarily defined
%% by their values; a general apparatus for computation-respecting refinements will
%% allow us to define, for instance, a type for partial functions.
%%
%% Then, for any refinement $\Refines{\Upsilon}{\phi}{\tau}$, we say
%% $\IsType{\Upsilon}{\phi}{\tau}$ just when for any
%% \begin{tikzcd}[cramped]
%%   \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
%% \end{tikzcd}%
%% , and for all $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$, from
%% $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$ we
%% can conclude $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
%% Stated internally as a presheaf in the topos, the meaning of parametric typehood for
%% $\phi$ is perhaps a bit more clear:
%% \[
%%   \Define{
%%     \IsType{-}{\phi}{\tau}
%%   }{
%%     \prod_{L,M,N\in\Exprs_\tau}
%%     {\phi(L,N)}^{L\sim_\tau M \times \phi(M,N)}
%%   }
%% \]
%%
%% \begin{thm}
%%   For any refinement $\Refines{\Upsilon}{\phi}{\tau}$, from
%%   $\IsValueType{\Upsilon}{\phi}{\tau}$ we may conclude
%%   $\IsType{\Upsilon}{\phi}{\tau}$.
%% \end{thm}
%% \begin{proof}
%%   Fix a renaming
%%   \begin{tikzcd}[cramped]
%%     \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
%%   \end{tikzcd}%
%%   , terms $\Member{L,M,N}{\Exprs_\tau(\Upsilon')}$ such that
%%   $\SBinRel{\sim_\tau}{\Upsilon'}{L}{M}$ and $\RMemEq{\Upsilon'}{M}{N}{\phi}$;
%%   by hypothesis, we also have some
%%   $\Of{\psi}{\Hom{\Sets^\SCtx}{\Yoneda{\Upsilon}}{\PERs{\Values_\tau^2}}}$,
%%   factoring through $\phi$.
%%
%%   We need to show that $\RMemEq{\Upsilon'}{L}{N}{\phi}$.
%%
%% \end{proof}
